Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-07-09 draft.tex
Make appendices more consistent

The class of $P\to P_1 P_2 \ell_1\bar\ell_2$ decays, where $P,P_1$, and $P_2$ are pseudoscalar mesons,
and $\ell_1$ and $\bar\ell_2$ are charged or neutral leptons, includes decays
such as $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ and $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$. \fred{If this appendix includes the previous one, how come there are less angular functions here?}
Their PDFs can generally be written as \cite{Lee:1992ih}
\begin{equation}
    P(q^2, k^2, \cos\theta_1, \cos \theta_2, \phi) = \frac{1}{\langle \Gamma\rangle}\frac{\dd^5 \Gamma}{\dd q^2\,\dd k^2\, \dd\cos\theta_1\,\dd\cos\theta_2\,\dd\phi} = \frac{1}{4\pi} \sum_i \frac{I_i(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\cos\theta_1,\phi)\,,
\end{equation}
where
\begin{equation}
\begin{aligned}
    f_1 & = 1\,,              &
    f_2 & = \cos 2\theta_1\,, &
    f_6 & = \cos  \theta_1\,, \\
    f_3 & = \sin^2\theta_1 \cos 2\phi\,,&
    f_4 & = \sin 2\theta_1 \cos  \phi\,,&
    f_5 & = \sin  \theta_1 \cos  \phi\,,\\
    f_9 & = \sin^2\theta_1 \sin 2\phi\,.&
    f_8 & = \sin 2\theta_1 \sin  \phi\,,&
    f_7 & = \sin  \theta_1 \sin  \phi\,,
\end{aligned}
\end{equation}
Here we denote dilepton mass squared and the dilepton helicity angle as $q^2$ and $\theta_1$, while
the dimeson mass squared and the dimeson helicity angles are denoted as $k^2$ and $\theta_2$. The
azimuthal angle is denoted as $\phi$. The total decay width is
\begin{equation}
    \langle \Gamma\rangle = \int_0^\pi \dd\theta_1\,\dd q^2\,\dd k^2\,\sin\theta_2 \frac{3I_1(q^2, k^2, \cos\theta_2) - I_2(q^2, k^2, \cos\theta_2)}{3}\,.
\end{equation}
As usual, we transform from a $\cos\theta_1$-dependence to a $\theta_1$-dependence.
\begin{equation}
\begin{aligned}
    P(q^2, k^2, \theta_1, \theta_2, \phi)
    & = \frac{1}{4\pi} \sum_i \frac{I(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\theta_1, \phi) \sin\theta_1 \sin\theta_2\\
    & \equiv \frac{1}{4\pi} \sum_i P_i(q^2, k^2, \theta_2) f_i(\theta_1, \phi)\,.
\end{aligned}
\end{equation}
where we also define $P_i(q^2, k^2, \theta_2) \equiv \sin\theta_2 I_i(q^2, k^2, \cos\theta_2) / \Gamma$.\\

We provide a basis $\hat{f}_i$ of angular functions which is dual to the basis $f_j$ in the sense that
\begin{equation}
    \frac{1}{4\pi}\int_0^\pi \dd\theta_1 \int_0^{2\pi}\dd \phi \hat{f}_i(\theta_1,\phi) f_j(\theta_1,\phi) \sin\theta_1 = \delta_{ij}\,,
\end{equation}
thereby enabling us to project out exactly one of the angular
observables $P_i(q^2, k^2, \theta_2)$
\begin{equation}
    \int_0^\pi \dd\theta_1 \int_0^{2\pi}\dd \phi \hat{f}_i(\theta_1,\phi) P(q^2, k^2, \theta_1,\theta_2,\phi) = P_i(q^2, k^2, \theta_2)\,.
\end{equation}
The dual basis reads
\begin{equation}
\begin{aligned}
    \hat{f}_1 & = \frac{21 + 15 \cos 2\theta_1}{16}\,, &
    \hat{f}_2 & = \frac{15 + 45 \cos 2\theta_1}{16}\,, &
    \hat{f}_6 & = 3 \cos\theta_1                   \,, \\
    \hat{f}_3 & = \frac{15}{4}                \cos 2\phi     \,, &
    \hat{f}_4 & = \frac{15}{4} \sin 2\theta_1 \cos  \phi     \,, &
    \hat{f}_5 & = 3            \sin  \theta_1 \cos  \phi     \,, \\
    \hat{f}_9 & = \frac{15}{4}                \sin 2\phi     \,, &
    \hat{f}_8 & = \frac{15}{4} \sin 2\theta_1 \sin  \phi     \,, &
    \hat{f}_7 & = 3            \sin  \theta_1 \sin  \phi     \,.
\end{aligned}
\end{equation}

@@

The physical PDF for the decay -- in the presence of Standard Model operator and their chirality flipped counter parts -- reads \cite{Boeer:2014xx}
\begin{equation}
    P_{\Lambda_b}(q^2, \cos\theta_1, \cos\theta_2, \phi) = \frac{1}{\Gamma} \frac{\dd^4\Gamma}{\dd q^2\,\dd \cos\theta_1\,\dd \cos\theta_2\,\dd \phi} = \frac{3}{8\pi} \sum_i \frac{K_{i}(q^2)}{\Gamma} f_i(\cos\theta_1, \cos\theta_2, \phi)\,,
\end{equation}
where $q^2$ denotes the dilepton mass squared, and $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and $N\pi$ systems, respectively. The basis of angular functions reads \fred{cite Danny's upcoming paper}
\begin{equation}
    \begin{aligned}
% 1
        f_{1ss} & = \sin^2\theta_1 &
        f_{1cc} & = \cos^2\theta_1 &
        f_{1c}  & = \cos\theta_1\\
% 2
        f_{2ss} & = \sin^2\theta_1 \cos\theta_2 &
        f_{2cc} & = \cos^2\theta_1 \cos\theta_2 &
        f_{2c}  & = \cos\theta_1   \cos\theta_2 \\
% 3
        f_{3sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\phi &
        f_{3s}  & = \sin\theta_1              \sin\theta_2 \cos\phi \\
% 4
        f_{4sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\phi &
        f_{4s}  & = \sin\theta_1              \sin\theta_2 \sin\phi
    \end{aligned}
\end{equation}
The decay width reads $\langle\Gamma\rangle = \langle2 K_{1ss} + K_{1cc}\rangle$. In terms of the angles $\theta_1$ and $\theta_2$, instead of the cosines, the PDF reads
\begin{equation}
    P_{\Lambda_b}(q^2, \theta_1, \theta_2, \phi) = \frac{3}{8\pi} \sum_i P_i(q^2) f_i(\cos\theta_1, \cos\theta_2, \phi) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $P_i(q^2) = K_i(q^2) / \Gamma$.

We provide a dual basis $\hat{f}_i$ of the angular functions which is orthogonal to the basis $f_j$ in the sense that
\begin{equation}
    \frac{3}{8\pi}\int_0^\pi \dd\theta_1 \int_0^\pi\,\dd\theta_2\, \int_0^{2\pi}\dd \phi \hat{f}_i(\theta_1,\theta_2,\phi) f_j(\theta_1,\theta_2,\phi) \sin\theta_1 \sin\theta_2 = \delta_{ij}\,,
\end{equation}
thereby enabling us to project out exactly one of the angular observables $P_i(q^2)$ using
\begin{equation}
    \int_0^\pi \dd \theta_1 \int_0^\pi \dd \theta_2 \int_0^{2\pi} \dd \phi P(q^2, \theta_1, \theta_2, \phi) \hat{f}_i(\theta_1, \theta_2, \phi) = P_i(q^2)\,.
\end{equation}
The dual basis reads
\begin{equation}
    \begin{aligned}
% 1
        \hat f_{1ss} & = \frac{\pi^3}{4}\big(3 \sin^2\theta_1 - 2 \cos^2\theta_1\big) &
        \hat f_{1cc} & = \frac{\pi^3}{2}\big(- \sin^2\theta_1 + 4 \cos^2\theta_1\big) &
        \hat f_{1c}  & = \pi^3 \cos\theta_1\\
% 2
        \hat f_{2ss} & = \frac{3\pi^3}{4}\big(3 \sin^2\theta_1 - 2 \cos^2\theta_1\big) \cos\theta_2 &
        \hat f_{2cc} & = \frac{3\pi^3}{2}\big(- \sin^2\theta_1 + 4 \cos^2\theta_1\big) \cos\theta_2 &
        \hat f_{2c}  & = 3\pi^3 \cos\theta_1   \cos\theta_2 \\
% 3
        \hat f_{3sc} & = \frac{15\pi^3}{2}\sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\phi &
        \hat f_{3s}  & = \frac{3\pi^3}{2} \sin\theta_1              \sin\theta_2 \cos\phi \\
% 4
        \hat f_{4sc} & = \frac{15\pi^3}{2}\sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\phi &
        \hat f_{4s}  & = \frac{3\pi^3}{2} \sin\theta_1              \sin\theta_2 \sin\phi
    \end{aligned}
\end{equation}

>>>>

The class of $P\to P_1 P_2 \ell_1\bar\ell_2$ decays, where $P,P_1$, and $P_2$ are pseudoscalar mesons,
and $\ell_1$ and $\bar\ell_2$ are charged or neutral leptons, includes decays
such as $\bar{B}\to\bar{K}\pi\ell^+\ell^-$ and $\bar{B}\to \pi\pi\ell^-\bar\nu_\ell$.
Their PDFs can generally be written as \cite{Lee:1992ih}
\begin{equation}
    P(q^2, k^2, \cos\theta_1, \cos \theta_2, \phi) = \frac{1}{\langle \Gamma\rangle}\frac{\dd^5 \Gamma}{\dd q^2\,\dd k^2\, \dd\cos\theta_1\,\dd\cos\theta_2\,\dd\phi} = \frac{1}{4\pi} \sum_i \frac{I_i(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\cos\theta_1,\phi)\,,
\end{equation}
where
\begin{equation}
\begin{aligned}
    f_1 & = 1\,,              &
    f_2 & = \cos 2\theta_1\,, &
    f_6 & = \cos  \theta_1\,, \\
    f_3 & = \sin^2\theta_1 \cos 2\phi\,,&
    f_4 & = \sin 2\theta_1 \cos  \phi\,,&
    f_5 & = \sin  \theta_1 \cos  \phi\,,\\
    f_9 & = \sin^2\theta_1 \sin 2\phi\,.&
    f_8 & = \sin 2\theta_1 \sin  \phi\,,&
    f_7 & = \sin  \theta_1 \sin  \phi\,,
\end{aligned}
\end{equation}
Here we denote dilepton mass squared and the dilepton helicity angle as $q^2$ and $\theta_1$, while
the dimeson mass squared and the dimeson helicity angles are denoted as $k^2$ and $\theta_2$. The
azimuthal angle is denoted as $\phi$. The total decay width is
\begin{equation}
    \langle \Gamma\rangle = \int_0^\pi \dd\theta_1\,\dd q^2\,\dd k^2\,\sin\theta_2 \frac{3I_1(q^2, k^2, \cos\theta_2) - I_2(q^2, k^2, \cos\theta_2)}{3}\,.
\end{equation}
As usual, we transform from a $\cos\theta_1$-dependence to a $\theta_1$-dependence.
\begin{equation}
\begin{aligned}
    P(q^2, k^2, \theta_1, \theta_2, \phi)
    & = \frac{1}{4\pi} \sum_i \frac{I(q^2, k^2, \cos\theta_2)}{\Gamma} f_i(\theta_1, \phi) \sin\theta_1 \sin\theta_2\\
    & \equiv \frac{1}{4\pi} \sum_i S_i(q^2, k^2, \theta_2) f_i(\theta_1, \phi)\,.
\end{aligned}
\end{equation}
where we also define $S_i(q^2, k^2, \theta_2) \equiv \sin\theta_2 I_i(q^2, k^2, \cos\theta_2) / (\dd\Gamma/\dd q^2)$.\\

The dual basis reads
\begin{equation}
\begin{aligned}
    \hat{f}_1 & = \frac{21 + 15 \cos 2\theta_1}{16}\,, &
    \hat{f}_2 & = \frac{15 + 45 \cos 2\theta_1}{16}\,, &
    \hat{f}_6 & = 3 \cos\theta_1                   \,, \\
    \hat{f}_3 & = \frac{15}{4}                \cos 2\phi     \,, &
    \hat{f}_4 & = \frac{15}{4} \sin 2\theta_1 \cos  \phi     \,, &
    \hat{f}_5 & = 3            \sin  \theta_1 \cos  \phi     \,, \\
    \hat{f}_9 & = \frac{15}{4}                \sin 2\phi     \,, &
    \hat{f}_8 & = \frac{15}{4} \sin 2\theta_1 \sin  \phi     \,, &
    \hat{f}_7 & = 3            \sin  \theta_1 \sin  \phi     \,.
\end{aligned}
\end{equation}
such that
\begin{equation}
    \int_0^\pi \dd\theta_1 \int_0^{2\pi}\dd \phi \hat{f}_i(\theta_1,\phi) P(q^2, k^2, \theta_1,\theta_2,\phi) = S_i(q^2, k^2, \theta_2)\,.
\end{equation}

@@

The physical PDF for the decay -- in the presence of Standard Model operator and their chirality flipped counter parts -- reads \cite{Boeer:2014xx}
\begin{equation}
    P_{\Lambda_b}(q^2, \cos\theta_1, \cos\theta_2, \phi) = \frac{1}{\dd\Gamma/\dd q^2} \frac{\dd^4\Gamma}{\dd q^2\,\dd \cos\theta_1\,\dd \cos\theta_2\,\dd \phi} = \frac{3}{8\pi} \sum_i \frac{K_{i}(q^2)}{\dd \Gamma/\dd q^2} f_i(\cos\theta_1, \cos\theta_2, \phi)\,,
\end{equation}
where $q^2$ denotes the dilepton mass squared, $\theta_1 \equiv \theta_\ell$ and $\theta_2 \equiv \theta_\Lambda$ denote the
helicity angles in the dilepton and $N\pi$ systems, respectively, and $\theta_3 = \phi$ denotes the azimuthal angle.
The basis of angular functions reads \cite{Boeer:2014xx}
\begin{equation}
    \begin{aligned}
% 1
        f_{1ss} & = \sin^2\theta_1 &
        f_{1cc} & = \cos^2\theta_1 &
        f_{1c}  & = \cos\theta_1\\
% 2
        f_{2ss} & = \sin^2\theta_1 \cos\theta_2 &
        f_{2cc} & = \cos^2\theta_1 \cos\theta_2 &
        f_{2c}  & = \cos\theta_1   \cos\theta_2 \\
% 3
        f_{3sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\phi &
        f_{3s}  & = \sin\theta_1              \sin\theta_2 \cos\phi \\
% 4
        f_{4sc} & = \sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\phi &
        f_{4s}  & = \sin\theta_1              \sin\theta_2 \sin\phi
    \end{aligned}
\end{equation}
The decay width reads
\begin{equation}
    \frac{\dd \Gamma}{\dd q^2} = 2 K_{1ss} + K_{1cc}
\end{equation}
In terms of the angles $\theta_1$ and $\theta_2$, instead of the cosines, the PDF reads
\begin{equation}
    P_{\Lambda_b}(q^2, \theta_1, \theta_2, \phi) = \frac{3}{8\pi} \sum_i S_i(q^2) f_i(\cos\theta_1, \cos\theta_2, \phi) \sin\theta_1 \sin\theta_2\,,
\end{equation}
where we also define $S_i(q^2) = K_i(q^2) / (\dd\Gamma/\dd q^2)$.\\

The dual basis reads
\begin{equation}
    \begin{aligned}
% 1
        \hat f_{1ss} & = \frac{\pi^3}{4}\big(3 \sin^2\theta_1 - 2 \cos^2\theta_1\big) &
        \hat f_{1cc} & = \frac{\pi^3}{2}\big(- \sin^2\theta_1 + 4 \cos^2\theta_1\big) &
        \hat f_{1c}  & = \pi^3 \cos\theta_1\\
% 2
        \hat f_{2ss} & = \frac{3\pi^3}{4}\big(3 \sin^2\theta_1 - 2 \cos^2\theta_1\big) \cos\theta_2 &
        \hat f_{2cc} & = \frac{3\pi^3}{2}\big(- \sin^2\theta_1 + 4 \cos^2\theta_1\big) \cos\theta_2 &
        \hat f_{2c}  & = 3\pi^3 \cos\theta_1   \cos\theta_2 \\
% 3
        \hat f_{3sc} & = \frac{15\pi^3}{2}\sin\theta_1 \cos\theta_1 \sin\theta_2 \cos\phi &
        \hat f_{3s}  & = \frac{3\pi^3}{2} \sin\theta_1              \sin\theta_2 \cos\phi \\
% 4
        \hat f_{4sc} & = \frac{15\pi^3}{2}\sin\theta_1 \cos\theta_1 \sin\theta_2 \sin\phi &
        \hat f_{4s}  & = \frac{3\pi^3}{2} \sin\theta_1              \sin\theta_2 \sin\phi
    \end{aligned}
\end{equation}
such that
\begin{equation}
    \int_0^\pi \dd \theta_1 \int_0^\pi \dd \theta_2 \int_0^{2\pi} \dd \phi P(q^2, \theta_1, \theta_2, \phi) \hat{f}_i(\theta_1, \theta_2, \phi) = S_i(q^2)\,.
\end{equation}
